---
title: "getchinson vs soil"
author: "GrGladkov"
date: "16 03 2020"
output: html_document
---
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1800)
```

```{r setup, include=FALSE}
setwd("~/storage/plates/plos/")

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
knitr::opts_knit$set(stop_on_error = 2L)
```

##Import old shady "plates" workspace, narrow down new_plates datacet, export into rds object, clear all shit and recreate workspace. Quite enought shamefull relocate working directory. 

```{r, message=FALSE, eval = FALSE}
load("~/storage/r-base_files/plates.RData")
library(phyloseq)
ps.f.plos <- prune_samples(sample_data(ps.f)$Medium %in% c("getchinson", "soil dna"),ps.f)
ps.f.plos <- prune_taxa(taxa_sums(ps.f.plos) > 0, ps.f.plos)
saveRDS(ps.f.plos, file = "~/storage/plates/ps.f.plos.rds")
```

```{r, message=FALSE}
setwd("~/storage/plates/plos/")
library(phyloseq)
ps <- readRDS(file = "~/storage/plates/ps.f.plos.rds")
ps
```

###Richness/depth correlation. Outliers - sample 81.
```{r, message=FALSE,echo=TRUE, fig.height=8, fig.width=12}
library(phyloseq)
library(ggplot2)
plot_rich_reads_samlenames_lm <- function(ps.f){
  require(phyloseq)
  rish <- phyloseq::estimate_richness(ps.f, measures = "Observed")
  reads.sum <- as.data.frame(sample_sums(ps.f))
  reads.summary <- cbind(rish, reads.sum)
  colnames(reads.summary) <- c("otus","reads")
  reads.summary["Description"] <- ps.f@sam_data$Number
  reads.summary["Repeats"] <- ps.f@sam_data$Repeats
  library(ggrepel)
  require(ggforce)
  p1 <- ggplot(data=reads.summary) + geom_point(aes(y=otus, x=log2(reads), color=Repeats),size=3) + geom_text_repel(aes(y=otus, x=log2(reads), label=Description)) + theme_bw()+
    geom_smooth(aes(y=otus, x=log2(reads), fill=Repeats, color=Repeats),method="lm", se=FALSE, ymin = 1)  + scale_x_continuous(sec.axis = sec_axis(sec.axis ~ 2**.))
  # geom_mark_ellipse(aes(y = otus, x=reads, group = Repeats, label = Repeats, color = Repeats), label.fontsize = 10, label.buffer = unit(2, "mm"), label.minwidth = unit(5, "mm"),con.cap = unit(0.1, "mm"))
  
  return(p1)
}
plot_rich_reads_samlenames_lm(ps)

```
###[Ricness/depth without outliers]
```{r, message=FALSE,echo=TRUE, fig.height=10, fig.width=14}
# ps.f <- prune_samples(sample_data(ps)$Number != c("81"), ps)
# 
# plot_rich_reads_samlenames_lm(ps.f)
ps.f <- ps
```
###
```{r}

```

###Split dataset by medium/soil

```{r, message=FALSE}

ps.f <- ps

ps.medium <- prune_samples(sample_data(ps.f)$Medium_or_Soil %in% c("medium"),ps.f)
ps.medium  <- prune_taxa(taxa_sums(ps.medium ) > 0, ps.medium )

ps.soil <- prune_samples(sample_data(ps.f)$Medium_or_Soil %in% c("soil"), ps.f)
ps.soil <- prune_taxa(taxa_sums(ps.soil) > 0, ps.soil)

# sod-podzolic - SP
# black soil - CZ

ps.cz <- prune_samples(sample_data(ps.f)$Soil %in% c("black soil"),ps.f)
ps.cz  <- prune_taxa(taxa_sums(ps.cz ) > 0, ps.cz )

ps.sp <- prune_samples(sample_data(ps.f)$Soil %in% c("sod-podzolic"), ps.f)
ps.sp <- prune_taxa(taxa_sums(ps.sp) > 0, ps.sp)

```

###Beta-diversity plots

```{r, message=FALSE,fig.height=8, fig.width=18, results=FALSE, fig.keep='all'}
beta_custom_norm_NMDS_elli <- function(ps, seed = 7888, normtype="vst", color="Repeats"){
  require(phyloseq)
  require(ggplot2)
  require(ggpubr)
  require(DESeq2)
  library(ggforce)
  
  
  # beta_NMDS <- function(){
  #normalisation. unifrac - rarefaction; wunifrac,bray - varstab
  
  diagdds = phyloseq_to_deseq2(ps, ~ Repeats)                  
  diagdds = estimateSizeFactors(diagdds, type="poscounts")
  diagdds = estimateDispersions(diagdds, fitType = "local") 
  if (normtype =="vst")
    pst <- varianceStabilizingTransformation(diagdds)
  if (normtype =="log") 
    pst <- rlogTransformation(diagdds)
  
  pst.dimmed <- t(as.matrix(assay(pst))) 
  pst.dimmed[pst.dimmed < 0.0] <- 0.0
  ps.varstab <- ps
  otu_table(ps.varstab) <- otu_table(pst.dimmed, taxa_are_rows = FALSE) 
  
  ps.rand <- rarefy_even_depth(ps, rngseed = seed)
  
  #beta and ordination
  
  ordination.b <- ordinate(ps.varstab, "NMDS", "bray")
  ordination.u <- ordinate(ps.rand, "NMDS", "unifrac")
  ordination.w <- ordinate(ps.varstab, "NMDS", "wunifrac")
  
  #plotting
  p1 = plot_ordination(ps, ordination.b, type="sample", color="Repeats", title="NMDS - Bray", 
                       axes = c(1,2) ) + theme_bw() + theme(text = element_text(size = 10)) + geom_point(size = 3) +
    geom_mark_ellipse(aes(group = Repeats, label = Repeats), label.fontsize = 10, label.buffer = unit(2, "mm"), label.minwidth = unit(5, "mm"),con.cap = unit(0.1, "mm"))
  
  p2 = plot_ordination(ps, ordination.u, type="sample", color="Repeats", title="NMDS - unifrac", 
                       axes = c(1,2) ) + theme_bw() + theme(text = element_text(size = 10)) + geom_point(size = 3) +
    geom_mark_ellipse(aes(group = Repeats, label = Repeats), label.fontsize = 10, label.buffer = unit(2, "mm"), label.minwidth = unit(5, "mm"),con.cap = unit(0.1, "mm"))
  
  p3 = plot_ordination(ps, ordination.w, type="sample", color="Repeats", title="NMDS - wunifrac", 
                       axes = c(1,2) ) + theme_bw() + theme(text = element_text(size = 10)) + geom_point(size = 3) + 
    geom_mark_ellipse(aes(group = Repeats, label = Repeats), label.fontsize = 10, label.buffer = unit(2, "mm"), label.minwidth = unit(5, "mm"),con.cap = unit(0.1, "mm"))
  
  #merge by ggpubr
  
  p.all <- ggarrange(p1, p2, p3, ncol = 3 , nrow = 1, common.legend = TRUE, legend = "none", font.label = list(size = 12, face = "bold", color ="black"))
  
  return(p.all)
}

beta_custom_norm_NMDS_elli(ps.f)
```

###Basic DESeq2 by Soil variable

```{r,message=FALSE}
Des.soil.w.simper <- function(ps){
  require(DESeq2)
  require(vegan)
  require(tibble)
  
  diagdds = phyloseq_to_deseq2(ps, ~ Soil)                  
  diagdds = estimateSizeFactors(diagdds, type="poscounts")
  diagdds = estimateDispersions(diagdds, fitType = "local") 
  diagdds = DESeq(diagdds)
  samp <-sample_data(ps)
  dds.counts <- diagdds@assays@.xData$data$counts
  dds.counts.df <- as.data.frame(dds.counts)
  aggdata <- t(aggregate.data.frame(as.data.frame(as.data.frame(t(diagdds@assays@.xData$data$mu))), by=list(samp$Soil), median))
  colnames(aggdata) <- aggdata[1,]
  aggdata <- aggdata[-1,]
  res = results(diagdds)
  res.df <- as.data.frame(res)
  nice <- cbind(res.df,as.data.frame(tax_table(ps)[rownames(res.df),]), as.data.frame(aggdata)[rownames(res.df),])               
  return(nice)
} 
des.soil <- Des.soil.w.simper(ps.soil) 
des.medium <- Des.soil.w.simper(ps.medium) 
```

####filtration ASVs in the DESeq2 result tables by abundanse and change confidence level

```{r,message=FALSE}
filter.des <- function(des, alpha=0.1, baseMean = 15){
  des <- des[(des$padj < alpha), ]
  des <- des[(des$baseMean > baseMean),]
  #des <- des[(des$log2FoldChange > 0),]
  des <-  des[(is.na(des$padj) != TRUE),]
  return(des)
}
alpha.soil <- 0.05
baseMean.soil <- 60
des.soil.filt <- filter.des(des.soil, alpha = alpha.soil, baseMean = baseMean.soil)
len.soil <- length(rownames(des.soil.filt))
paste0("length des.soil after filtation with param: alpha - ", alpha.soil, ", baseMean - ", baseMean.soil, " ==> ", len.soil)

alpha.medium <- 0.05
baseMean.medium <- 60
des.medium.filt <- filter.des(des.medium, alpha = alpha.medium, baseMean = baseMean.medium)
len.medium <- length(rownames(des.medium.filt))
paste0("length des.medium  after filtation with param: alpha - ", alpha.soil, ", baseMean - ", baseMean.soil, " ==> ", len.medium)

```

####Tree for soil soil-specific taxa

```{r,  message=FALSE, results=FALSE, echo=FALSE, fig.height=12, fig.width=16}
list.both <- c(rownames(des.soil.filt), rownames(des.medium.filt)) 
ps.changed <- prune_taxa(list.both, ps.f)
phyloseq <- ps.changed
library(ggtree)
library(tidyverse)
tree <- phyloseq@phy_tree
taxa.pruned <- as.data.frame(phyloseq@tax_table@.Data)
taxa.pruned <- taxa.pruned %>%  mutate_all(as.character)
old.tips <- tree$tip.label
matches <- regmatches(tree$tip.label, gregexpr("[[:digit:]]+", tree$tip.label))
taxa.pruned$number <- as.numeric(unlist(matches))
taxa.pruned$taxa <- ifelse(is.na(taxa.pruned$Genus), taxa.pruned$Family, taxa.pruned$Genus)
taxa.pruned[taxa.pruned == "Burkholderia-Caballeronia-Paraburkholderia"] <- "Burkholderia"
taxa.pruned[taxa.pruned == "Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium"] <- "Pararhizobium"
taxa.pruned$taxa2 <- ifelse(is.na(taxa.pruned$Species), with(taxa.pruned, paste0(taxa)), with(taxa.pruned, paste0(taxa, " ", Species )))
taxa.pruned$taxa3 <- ifelse(taxa.pruned$Phylum == "Proteobacteria", with(taxa.pruned, paste0(taxa.pruned$number, ".", taxa2, " // ", Class)), with(taxa.pruned, paste0(taxa.pruned$number, ".", taxa2, " // ", Phylum)))
class(taxa.pruned$Kingdom)
tree$tip.label <- taxa.pruned$taxa3
p <- ggtree(tree, ladderize = F) + geom_tiplab(mapping = aes(), align=TRUE, linesize=.5) + xlim(NA, 4)
p
```
####Add log2FoldChange for tree. Red - prevalence in CZ, blue - in SP.

```{r,  message=FALSE,echo=TRUE, fig.height=12, fig.width=14}
t.medium <-   as_tibble(des.medium, rownames = "id") %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>% 
  select(id, log2FoldChange) %>%  
  rename(medium = log2FoldChange)

t.soil <- as_tibble(des.soil, rownames = "id") %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>% 
  select(id, log2FoldChange) %>%  
  rename(soil = log2FoldChange)

t.all <- full_join(t.medium, t.soil)

names.d <-  cbind(long = tree$tip.label, short = old.tips) %>% data.frame(row.names = "short")
t.names <- as_tibble(names.d, rownames = "id")
d.all.tips <- full_join(t.names, t.all) %>%  select(-c(id)) %>% 
  as.data.frame() %>% column_to_rownames("long")

p2 <- gheatmap(p, d.all.tips, offset=2, width=0.8)  + scale_colour_gradientn(colours = c("blue", "white", "red"),na.value = "grey50", guide = "colourbar",aesthetics = "fill") 
p2 <- p2 + labs(fill = "L2FC CZ/SP")
```

```{r, message=FALSE,echo=TRUE, fig.height=14, fig.width=14}
otus.medium.rel <- t(apply(otu_table(ps.medium), 1, function(x) x / sum(x)))
otus.soil.rel <- t(apply(otu_table(ps.soil), 1, function(x) x / sum(x)))
# for medium
as_tibble(t(otus.medium.rel ), rownames = "id" ) %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>%
  select(id) -> id.medium
as_tibble(t(otus.medium.rel), rownames = "id" ) %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>%
  select_if(is.numeric) %>% 
  replace(is.na(.), 0) %>% 
  mutate(mediumMean=rowMeans(.)) %>% 
  select(mediumMean) -> res.medium
res.mediumMean.rel <- cbind2(id.medium,res.medium)

# for soil
as_tibble(t(otus.soil.rel), rownames = "id" ) %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>%
  select(id) -> id.soil
as_tibble(t(otus.soil.rel), rownames = "id" ) %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>%
  select_if(is.numeric) %>% 
  replace(is.na(.), 0) %>% 
  mutate(soilMean=rowMeans(.)) %>% 
  select(soilMean) -> res.soil
res.soilMean.rel <- cbind2(id.soil, res.soil)


all.resMean.rel <- full_join(res.mediumMean.rel, res.soilMean.rel)
all.resMean.rel %>% replace(is.na(.), 0) -> all.resMean.rel
all.resMean.rel$mediumMean <- 0 - all.resMean.rel$mediumMean
all.resMean.rel.melted <- reshape2::melt(all.resMean.rel, id=c("id"))

library(reshape2)
library(ggstance)
all.resMean.rel.melted <- melt(all.resMean.rel, id=c("id"))
first.tree <- phyloseq@phy_tree
ids.tree <-  data.frame(id = first.tree$tip.label, id.taxa = tree$tip.label)
some.join <- full_join(ids.tree, all.resMean.rel.melted)
some.join <- subset(some.join, select = -c(id))
some.join["numbers"] <-  ifelse(some.join$value == 0, NA, abs(some.join$value))
some.join$value <- some.join$value*100
some.join["position"] <-  ifelse(some.join$value >= 0, some.join$value + 0.5, some.join$value - 0.5)
p4 <- facet_plot(p2, panel = "medium / soil", data = some.join,  geom = geom_barh, 
                 mapping = aes(x = value, label=value), color = "salmon", fill = "white",
                 stat='identity')

p4 <- facet_plot(p4, geom=geom_text , data = some.join, mapping=aes(x=position,label=round(numbers*100, digits = 2)), panel = "medium / soil")

library(gtable)
library(grid)
gt <-  ggplot_gtable(ggplot_build(p4))
gt$widths[7] <-  0.5*gt$widths[7] # in this case it was colmun 7 - reduce the width by a half
plot(gt) # plot with grid draw

```

### DESeq between medium. Splitting factor - soil type. 
####Improve on des function - add a required factor(with as.formula function)

```{r, message=FALSE}
des_w_simper <- function(ps, factor){
  require(DESeq2)
  require(vegan)
  require(tibble)
  
  diagdds = phyloseq_to_deseq2(ps, as.formula(paste( "~", factor)))                  
  diagdds = estimateSizeFactors(diagdds, type="poscounts")
  diagdds = estimateDispersions(diagdds, fitType = "local") 
  diagdds = DESeq(diagdds)
  samp <-sample_data(ps)
  dds.counts <- diagdds@assays@.xData$data$counts
  dds.counts.df <- as.data.frame(dds.counts)
  aggdata <- t(aggregate.data.frame(as.data.frame(as.data.frame(t(diagdds@assays@.xData$data$mu))), by=list(samp[[factor]]), median))
  colnames(aggdata) <- aggdata[1,]
  aggdata <- aggdata[-1,]
  res = results(diagdds)
  res.df <- as.data.frame(res)
  nice <- cbind(res.df, as.data.frame(tax_table(ps)[rownames(res.df),]), as.data.frame(aggdata)[rownames(res.df),])               
  return(nice)
} 

des.sp <- des_w_simper(ps.sp, "Medium")
des.cz <- des_w_simper(ps.cz, "Medium")
des.cz

```

###[Fun for tree des visualisation]
####[Optimise tree size. Plot for Mean/padj selection] 

```{r, eval=FALSE}
 filter.des <- function(des, alpha=0.1, baseMean = 15){
  des <- des[(des$padj < alpha), ]
  des <- des[(des$baseMean > baseMean),]
  #des <- des[(des$log2FoldChange > 0),]
  des <-  des[(is.na(des$padj) != TRUE),]
  return(des)
}
alpha.soil <- 0.05
baseMean.soil <- 60
des.soil.filt <- filter.des(des.soil, alpha = alpha.soil, baseMean = baseMean.soil)
len.soil <- length(rownames(des.soil.filt))
paste0("length des.soil after filtation with param: alpha - ", alpha.soil, ", baseMean - ", baseMean.soil, " ==> ", len.soil)

seq.list <- seq(0, 500, by=10)
jac.steppo <- function(some_number){
  some_ps <- prune_taxa(taxa_sums(some_ps) > some_number , some_ps) 
  otus.dt <- as.data.frame(some_ps@otu_table@.Data)
  jacco.mean <- mean(vegdist(otus.dt, method = "jaccard"))
  return(jacco.mean)
}

some_ps <- ps.s
seq.list <- seq(0, 5000, by=10)
jacco.list <- sapply(seq.list, jac.steppo)
plot(jacco.list)
```


###UNIFRAC visualisation with step by step tree glom. Rarefaction.
####[draft]

```{r, fig.height=8, fig.width=12, eval=FALSE}

ps.soil@sam_data
seq.list <- seq(0, 0.8, by=0.005)
library(phyloseq)
library(future.apply)

uni.steppo.mean.soil <- function(some_number){
  ps.soil <- tip_glom(ps.soil, some_number) 
  jacco.mean <- mean(phyloseq::distance(ps.soil, method = "unifrac", type = "samples"))
  return(jacco.mean)
}

uni.steppo.mean.medium <- function(some_number){
  ps.medium <- tip_glom(ps.medium, some_number) 
  jacco.mean <- mean(phyloseq::distance(ps.medium, method = "unifrac", type = "samples"))
  return(jacco.mean)
}

uni.steppo.centroid.medium <- function(some_number){
  dist <- phyloseq::distance(ps.soil.rand, method = "unifrac", type = "samples")
  names.sp <- subset(ps.soil.rand@sam_data, ps.soil.rand@sam_data$Soil %in% "sod-podzolic") %>% rownames()
  names.cz <- subset(ps.soil.rand@sam_data, ps.soil.rand@sam_data$Soil %in% "black soil") %>% rownames()
  jacco.mean <- usedist::dist_between_centroids(dist, names.sp, names.cz, squared = TRUE)
  return(jacco.mean)
}

uni.steppo.centroid.medium <- function(some_number){
  dist <- phyloseq::distance(ps.soil.rand, method = "unifrac", type = "samples")
  names.sp <- subset(ps.soil.rand@sam_data, ps.soil.rand@sam_data$Soil %in% "sod-podzolic") %>% rownames()
  names.cz <- subset(ps.soil.rand@sam_data, ps.soil.rand@sam_data$Soil %in% "black soil") %>% rownames()
  jacco.mean <- usedist::dist_between_centroids(dist, names.sp, names.cz, squared = TRUE)
  return(jacco.mean)
}


plan(multiprocess, workers = 20)

phyloseq::distance(ps.soil, method = "unifrac", type = "samples")

uni.list.medium <- future_sapply(seq.list, uni.steppo.medium)
uni.list.soil <- future_sapply(seq.list, uni.steppo.soil)
uni.list.soil
future:::ClusterRegistry("stop")
uni.list.medium
df <-as.data.frame(cbind(seq.list, uni.list.medium, uni.list.soil), row.names = seq.list)
df
df.melted <- reshape2::melt(df, id=c("seq.list"))
df.melted
library(ggplot2)
p1 <- ggplot(data=df.melted) + geom_point(aes(y=value, x=seq.list, color=variable),size=3) 
p1


dis.medium <- phyloseq::distance(ps.medium, method = "unifrac", type = "samples")
dis.medium


veganifyOTU <- function(physeq){
  require(phyloseq)
  if(taxa_are_rows(physeq)){physeq <- t(physeq)}
  return(as(otu_table(physeq), "matrix"))
}

veg.medium <- veganifyOTU(ps.medium)
veg.soil <- veganifyOTU(ps.soil)

groups.medium <- ps.medium@sam_data$Soil
groups.soil <- ps.soil@sam_data$Soil

plan(multiprocess, workers = 20)

uni.steppo.betadisp.soil <- function(some_number, ps){
  ps.soil <- tip_glom(ps.soil, some_number) 
  dist <- phyloseq::distance(ps.soil, method = "unifrac", type = "samples")
  groups <- ps.soil@sam_data$Soil
  mod <- vegan::betadisper(dist, groups)
  return(jacco.mean)
}

uni.steppo.medium <- function(some_number){
  ps.medium <- tip_glom(ps.medium, some_number) 
  jacco.mean <- mean(phyloseq::distance(ps.medium, method = "unifrac", type = "samples"))
  return(jacco.mean)
}

?usedist::dist_between_centroids
future:::ClusterRegistry("stop")

ps.soil.rand <- rarefy_even_depth(ps.soil, rngseed = 7788)

dist <- phyloseq::distance(ps.soil.rand, method = "unifrac", type = "samples")
names.sp <- subset(ps.soil.rand@sam_data, ps.soil.rand@sam_data$Soil %in% "sod-podzolic") %>% rownames()
names.cz <- subset(ps.soil.rand@sam_data, ps.soil.rand@sam_data$Soil %in% "black soil") %>% rownames()
usedist::dist_between_centroids(dist, names.sp, names.cz, squared = TRUE)

groups <- ps.soil@sam_data$Soil
mod <- vegan::betadisper(dist, groups)
mod.box <- boxplot(mod)

mod

mod.box$stats
?boxplot
TukeyHSD(mod, which = "group", ordered = FALSE, conf.level = 0.95)
scores(mod, display = c("sites", "centroids"), choices = c(1,2))


?adonis2
```
####clean copy
```{r, fig.height=8, fig.width=12}

library(future)
library(future.apply)

plan(multiprocess, workers = 20)
seq.list <- seq(0, 0.6, by=0.005)
# seq.list <- seq(0, 0.6, by=0.1)

steppo.mean <- function(physeq, seq.list, rar = FALSE){
  
  message("steppo.mean run")
  
  if (rar) {
    physeq <- rarefy_even_depth(physeq,  rngseed = 7788, verbose = FALSE)
  } else {
    physeq <- physeq
  }
  
  uni.steppo.mean <- function(some_number){
    physeq.glom <- tip_glom(physeq, some_number) 
    jacco.mean <- mean(phyloseq::distance(physeq.glom, method = "unifrac", type = "samples"))
    return(jacco.mean)
  }
  
  uni.list <- future_sapply(seq.list, uni.steppo.mean)
  df <-as.data.frame(cbind(seq.list, uni.list), row.names = seq.list)
  message("OUT:")
  print.data.frame(df)
  return(df)
}

steppo.centroids <- function(physeq, seq.list, rar = FALSE){
  
  message("steppo.centroids run")
  if (rar) {
    physeq <- rarefy_even_depth(physeq,  rngseed = 7788, verbose = FALSE)
  } else {
    physeq <- physeq
  }
  
  uni.steppo.centroid <- function(some_number){
    ps.glommed <- tip_glom(physeq, some_number)
    dist <- phyloseq::distance(ps.glommed, method = "unifrac", type = "samples")
    names.sp <- subset(physeq@sam_data, physeq@sam_data$Soil %in% "sod-podzolic") %>% rownames()
    names.cz <- subset(physeq@sam_data, physeq@sam_data$Soil %in% "black soil") %>% rownames()
    jacco.mean <- usedist::dist_between_centroids(dist, names.sp, names.cz, squared = TRUE)
    return(jacco.mean)
  }

  uni.list <- future_sapply(seq.list, uni.steppo.centroid)
  df <-as.data.frame(cbind(seq.list, uni.list), row.names = seq.list)
  message("OUT:")
  print.data.frame(df)
  return(df)
}



steppo.plot <- function(ps.soil, ps.medium, steppo, seq.list, rar = FALSE){
  df.soil <- steppo(ps.soil, seq.list, rar)
  df.medium <- steppo(ps.medium, seq.list, rar)
  df <- merge(df.soil, df.medium, by="seq.list")
  colnames(df) <- c("seq.list", "soil", "medium")
  df.melted <- reshape2::melt(df, id=c("seq.list"))
  require(ggplot2)
  print.data.frame(df.melted)
  p <- ggplot(data=df.melted) + geom_point(aes(y=value, x=seq.list, color=variable),size=3) + theme_bw() 
  return(p)
}

p.mean.norar <- steppo.plot(ps.soil, ps.medium, steppo.mean , seq.list, rar = FALSE)
p.mean.rar <- steppo.plot(ps.soil, ps.medium, steppo.mean , seq.list, rar = TRUE)
p.centroid.norar <- steppo.plot(ps.soil, ps.medium, steppo.centroids , seq.list, rar = FALSE)
p.centroid.rar <- steppo.plot(ps.soil, ps.medium, steppo.centroids , seq.list, rar = TRUE)

p.mean.norar
p.mean.rar

library(ggpubr)
p <- ggarrange( p.mean.norar, p.mean.rar, p.centroid.norar, p.centroid.rar, nrow = 2, ncol = 2,
                labels=c("mean_rarefied", "mean_w_rarefaction", "centroids_rarefied", "centroids_w_rarefaction"), common.legend = TRUE)

p

future:::ClusterRegistry("stop")
```

### REWRITE! Not have enough time.  

```{r, message=FALSE}
alpha.sp <- 0.05
baseMean.sp <- 80
des.sp.filt <- filter.des(des.sp, alpha = alpha.sp, baseMean = baseMean.sp)
len.sp <- length(rownames(des.sp.filt))
# in fun
paste0("length des.sp after filtation with param: alpha - ", alpha.sp, ", baseMean - ", baseMean.sp, " ==> ", len.sp)

alpha.cz <- 0.05
baseMean.cz <- 80
des.cz.filt <- filter.des(des.cz, alpha = alpha.cz, baseMean = baseMean.cz)
len.cz <- length(rownames(des.cz.filt))
paste0("length des.cz after filtation with param: alpha - ", alpha.cz, ", baseMean - ", baseMean.cz, " ==> ", len.cz)

```

####REWRITE! Collect all eggs in one bucket

```{r,  message=FALSE, results=FALSE, echo=FALSE, fig.height=12, fig.width=16, eval=TRUE}
list.both <- c(rownames(des.sp.filt), rownames(des.cz.filt)) 
ps.changed <- prune_taxa(list.both, ps.f)
phyloseq <- ps.changed
library(ggtree)
library(tidyverse)
tree <- phyloseq@phy_tree
taxa.pruned <- as.data.frame(phyloseq@tax_table@.Data)
taxa.pruned <- taxa.pruned %>%  mutate_all(as.character)
old.tips <- tree$tip.label
matches <- regmatches(tree$tip.label, gregexpr("[[:digit:]]+", tree$tip.label))
taxa.pruned$number <- as.numeric(unlist(matches))
taxa.pruned$taxa <- ifelse(is.na(taxa.pruned$Genus), taxa.pruned$Family, taxa.pruned$Genus)
taxa.pruned[taxa.pruned == "Burkholderia-Caballeronia-Paraburkholderia"] <- "Burkholderia"
taxa.pruned[taxa.pruned == "Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium"] <- "Pararhizobium"
taxa.pruned$taxa2 <- ifelse(is.na(taxa.pruned$Species), with(taxa.pruned, paste0(taxa)), with(taxa.pruned, paste0(taxa, " ", Species )))
taxa.pruned$taxa3 <- ifelse(taxa.pruned$Phylum == "Proteobacteria", with(taxa.pruned, paste0(taxa.pruned$number, ".", taxa2, " // ", Class)), with(taxa.pruned, paste0(taxa.pruned$number, ".", taxa2, " // ", Phylum)))
class(taxa.pruned$Kingdom)
tree$tip.label <- taxa.pruned$taxa3
p <- ggtree(tree, ladderize = F) + geom_tiplab(mapping = aes(), align=TRUE, linesize=.5) + xlim(NA, 4)
t.cz <-   as_tibble(des.cz, rownames = "id") %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>% 
  select(id, log2FoldChange) %>%  
  rename(Getchinson = log2FoldChange)

t.sp <- as_tibble(des.sp, rownames = "id") %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>% 
  select(id, log2FoldChange) %>%  
  rename(Soil_DNA = log2FoldChange)

t.all <- full_join(t.cz, t.sp)

names.d <-  cbind(long = tree$tip.label, short = old.tips) %>% data.frame(row.names = "short")
t.names <- as_tibble(names.d, rownames = "id")
d.all.tips <- full_join(t.names, t.all) %>%  select(-c(id)) %>% 
  as.data.frame() %>% column_to_rownames("long")

p2 <- gheatmap(p, d.all.tips, offset=2, width=0.8)  + scale_colour_gradientn(colours = c("blue", "white", "red"),na.value = "grey50", guide = "colourbar",aesthetics = "fill") 
p2 <- p2 + labs(fill = "L2FC Soil_DNA/Getchinson")

otus.cz.rel <- t(apply(otu_table(ps.cz), 1, function(x) x / sum(x)))
otus.sp.rel <- t(apply(otu_table(ps.sp), 1, function(x) x / sum(x)))
# for cz
as_tibble(t(otus.cz.rel ), rownames = "id" ) %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>%
  select(id) -> id.cz
as_tibble(t(otus.cz.rel), rownames = "id" ) %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>%
  select_if(is.numeric) %>% 
  replace(is.na(.), 0) %>% 
  mutate(czMean=rowMeans(.)) %>% 
  select(czMean) -> res.cz
res.czMean.rel <- cbind2(id.cz,res.cz)

# for sp
as_tibble(t(otus.sp.rel), rownames = "id" ) %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>%
  select(id) -> id.sp
as_tibble(t(otus.sp.rel), rownames = "id" ) %>% 
  filter(id %in% phyloseq@phy_tree$tip.label) %>%
  select_if(is.numeric) %>% 
  replace(is.na(.), 0) %>% 
  mutate(spMean=rowMeans(.)) %>% 
  select(spMean) -> res.sp
res.spMean.rel <- cbind2(id.sp, res.sp)


all.resMean.rel <- full_join(res.czMean.rel, res.spMean.rel)
all.resMean.rel %>% replace(is.na(.), 0) -> all.resMean.rel
all.resMean.rel$czMean <- 0 - all.resMean.rel$czMean
all.resMean.rel.melted <- reshape2::melt(all.resMean.rel, id=c("id"))

library(reshape2)
library(ggstance)
all.resMean.rel.melted <- melt(all.resMean.rel, id=c("id"))
first.tree <- phyloseq@phy_tree
ids.tree <-  data.frame(id = first.tree$tip.label, id.taxa = tree$tip.label)
some.join <- full_join(ids.tree, all.resMean.rel.melted)
some.join <- subset(some.join, select = -c(id))
some.join["numbers"] <-  ifelse(some.join$value == 0, NA, abs(some.join$value))
some.join$value <- some.join$value*100
some.join["position"] <-  ifelse(some.join$value >= 0, some.join$value + 0.5, some.join$value - 0.5)
p4 <- facet_plot(p2, panel = "Soil_DNA / Getchinson", data = some.join,  geom = geom_barh, 
                 mapping = aes(x = value, label=value), color = "salmon", fill = "white",
                 stat='identity')

p4 <- facet_plot(p4, geom=geom_text , data = some.join, mapping=aes(x=position,label=round(numbers*100, digits = 2)), panel = "Soil_DNA / Getchinson")

library(gtable)
library(grid)
gt <-  ggplot_gtable(ggplot_build(p4))
gt$widths[7] <-  0.5*gt$widths[7] # in this case it was colmun 7 - reduce the width by a half
plot(gt) # plot with grid draw

```



##Add soil extracts plates to dataset

```{r,message = FALSE, , eval = FALSE}
load("~/storage/r-base_files/plates.RData")
library(phyloseq)
ps.f.plosmore <- prune_samples(sample_data(ps.f)$Medium %in% c("getchinson", "soil dna", "black soil extract", "sod-podzolic extract"),ps.f)
ps.f.plosmore <- prune_taxa(taxa_sums(ps.f.plosmore) > 0, ps.f.plosmore)
saveRDS(ps.f.plosmore, file = "~/storage/plates/ps.f.plosmore.rds")
quit(save = "no", status = 0, runLast = TRUE)
```

```{r}
samp <-sample_data(ps)
list(samp[["Soil"]])
```

```{r,message = FALSE}
setwd("~/storage/plates/plos/")
library(phyloseq)
ps.more <- readRDS(file = "~/storage/plates/ps.f.plosmore.rds")
ps.more@sam_data
```

###Otus/reads plot for expanded dataset

```{r, message=FALSE,echo=TRUE, fig.height=8, fig.width=12}
plot_rich_reads_samlenames_lm(ps.more)
```

###Beta plots with soil extracts

```{r, message=FALSE,echo=TRUE, fig.height=8, fig.width=16}
beta_custom_norm_NMDS_elli(ps.more)
```
